---
title: "Preparation for Working with Scholarly Literature in R: Pulling, Wrangling, Cleaning, and Analyzing Structured Bibliographic Metadata"
author: "Clarke Iakovakis"
date: "August 5-9, 2019"
output: 
  pdf_document:
  number_sections: TRUE
  theme: architect
  df_print: paged
  toc: true
  toc_depth: 3
header-includes:
  - \usepackage{xcolor}
  - \usepackage{fancyhdr}
urlcolor: blue
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(png)
```

# Introduction

Hello everyone. I'm looking forward to our AM4 sessions at [FSCI 2019](https://www.force11.org/fsci/2019) on ["AM4 - Working with Scholarly Literature in R: Pulling, Wrangling, Cleaning, and Analyzing Structured Bibliographic Metadata."](https://www.force11.org/fsci/2019/course-abstracts#AM4). I am [Clarke Iakovakis](https://info.library.okstate.edu/clarke-iakovakis), the Scholarly Services Librarian at Oklahoma State University. 

The workshop materials are available on the Google Drive folder for our course at http://tiny.cc/fsci19_am4. You can also find them [on my GitHub](https://github.com/ciakovx/rorcid.workshop). 

# Learning Objectives

A participant in this course will at the end of the course be able to:
* Use R to complete a range of data tasks, including navigating the RStudio environment, understanding R syntax and data types, installing and loading packages, reading data from files and APIs, and executing functions for tidying and  analyzing data,
* Use the rorcid package in R to extract data from ORCID profiles, including authenticating with the ORCID API, finding ORCID iDs by name/affiliation, getting author biographical information, and getting works associated with individuals.
* Use the rcrossref and roadoi packages in R to extract data from the Crossref and Unpaywall database, including authenticating with the Crossref API, forming queries, getting publication information, getting data on articles, converting articles to BibTeX/RIS, getting citation data, and checking for open access versions.


# Schedule

## Monday:
* Introduction to R/RStudio and its advantages over other data tools. Setting up R and RStudio and getting to know the interface.
* Nuts and bolts of R syntax: writing and evaluating expressions, assigning values, calling functions, understanding data types, reading data into R, working with data frames.
* Using dplyr functions for data manipulation: renaming/recoding/selecting/creating variables, recoding values, ordering data.

## Tuesday:


Thursday:

Friday:

# Set-up Instructions

I would ask that you complete the following if you have time before our session. Please [contact me](https://info.library.okstate.edu/clarke-iakovakis) if you have problems with any of this. You will be receiving another set of documents and code for working with the code packages during the session. This document is only to help you get set up.

# Download R & R Studio

To download R, go to <https://www.r-project.org/>. Click on CRAN (Comprehensive R Archive Network) under Download, and scroll down to your country. Select the download link corresponding to the city that is geographically closest to you.

Go to <https://www.rstudio.com/products/RStudio/#Desktop> to download the RStudio desktop software. RStudio is a user interface for working with R. It is called an Integrated Development Environment (IDE) and acts as a sort of wrapper around the R language. You can use R without RStudio, but it's much more limiting. RStudio makes it easier to import datasets, create and write scripts, and has an autocomplete activated for functions and variables you've already assigned. RStudio makes using R much more effective, and is also free and open source.

# Install R packages

Open R Studio and go to **File > Open File**. Navigate to the directory for the folder this document was in, and open **setup_code**.

Click on the first line, `install.packages("rorcid")`. Then click the **Run** button in the upper right corner of that window.

It will print some lines of code and make take a few seconds. If successful, it will tell you the packaged was successfully unpacked. Now run the next line, `library(rorcid)`. As long as it doesn't say "there is no package called 'rorcid', you should be good.

Do the same thing with the remaining packages.

# Set up `rorcid`

Next, you need to authenticate with an ORCID API Key. According to the [ORCID API tutorial](https://members.orcid.org/api/tutorial/read-orcid-records), anyone can receive a key to access the public API. 

Run the line `orcid_auth()` from the **setup_code** script. You should see a message stating: `no ORCID token found; attempting OAuth authentication` and a window will open in your default internet browser. Log-in to your orcid account. You will be asked to give `rorcid` authorization to access your ORCID Record for the purposes of getting your ORCID iD. Click "Authorize."

If successful, the browser window will state: "Authentication complete. Please close this page and return to R." Return to R Studio and you should see in your R console the word **Bearer**, followed by a long string of letters and numbers. These letters and numbers are your API key. At this point, this should be cached locally in your working directory. 

Highlight and copy the API key (the letters and numbers only--exclude the word "Bearer" and the space). Paste it into the quotation marks on the **setup_code** script in the line `ORCID_TOKEN="copy and paste your token here"`. Copy this line to the clipboard.

Click on the line that says `usethis::edit_r_environ()` and click the Run button. A new window will open in R Studio. Paste the line into that window. Leave the tab open. 

Navigate back to the **setup_code** and run the next `orcid_auth()` line. It should print the token to your R console.

# Set up `rcrossref`

In the **setup_code** script, replace your email in `crossref_email=name@example.com`. Copy this line to the clipboard.

Go back to the window that opened when you ran `edit_r_environ()` and paste in the `crossref_email`.

**Then press enter to create a new line below `crossref_email=name@example.com`, and leave it blank.** 

Your R environment should look like this

```{r,fig.height=1,echo=FALSE}
renviron <- readPNG("./r_environ.png")
grid::grid.raster(renviron)
```

Click **File > Save**. Then in the R Studio navigation menu, click Session > Restart R.

# Download files

Go to the Google Drive folder for our course at http://tiny.cc/fsci19_am4. Right click on the **Workshop Materials** folder and click Download. Save the folder to your computer.

# Help with R and R Studio

* `swirl` is a package you can install in R to learn about R and data science interactively. Just type `install.packages("swirl")` into your R console, load the package by typing `library("swirl")`, and then type `swirl()`. Read more at <http://swirlstats.com/>.
* [R For Data Science](http://r4ds.had.co.nz/) by Garrett Grolemund & Hadley Wickham is the best book out there for learning R.
* [Introduction to R for Libraries](https://doi.org/10.6084/m9.figshare.8029301) is a webinar recording, slides, handouts
* [R cheat sheets (all)](https://github.com/rstudio/cheatsheets)
  - [base R cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/base-r.pdf)
  - [purrr cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf)
  - [data transformation cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf)
  - [data import cheat sheet](https://github.com/rstudio/cheatsheets/blob/master/data-import.pdf)

